---
<<<<<<< HEAD
=======
title: "Reproducible Research: Peer Assessment 1"
>>>>>>> 80edf39c3bb508fee88e3394542f967dd3fd3270
output: 
  html_document:
    keep_md: true
---
<<<<<<< HEAD
#                Reproducible Research - RMarkdown/Knitr Project

## Libraries needed
```{r}
library(dplyr)
library(lubridate)
library(ggplot2)
```
myActivityFn <- function() {

## 1. Read the data file
``` {r}
activityData = read.csv("activity.csv", header=TRUE)
```

# --- Computation with original dataset ---

#### 2.a. Calculate the Mean, Total Steps for each day
``` {r}
meanPerDay = activityData %>% na.omit() %>% group_by(ymd(date)) %>% summarize(meanSteps=mean(steps), totalSteps = sum(steps))
```

#### 2.b. Print the calculated data and plot the Total Steps per Day
``` {r}
print(meanPerDay)
g= ggplot(meanPerDay, aes(totalSteps)) + geom_histogram(fill="green", binwidth=1000) + xlab("Steps") + ylab("Frequency") + ggtitle("Steps taken daily")
print(g)
```
``` {r, echo=FALSE}
png("MeanStepsPerDay.png")
dev.set(2)
dev.copy(png, "Figure/MeanStepsPerDay.png")
dev.off()
```

#### 2.c. Calculate the Mean and Median of the Total steps taken
``` {r}
meanAndMedian = meanPerDay %>% na.omit() %>% summarize(MeanofTotalSteps = mean(totalSteps), MedianofTotalSteps = median(totalSteps))
print(meanAndMedian)
```

#### 3.a. Calculate the AVergae Daily Activity (Steps)
``` {r}
meanPerInterval = activityData %>% na.omit() %>% group_by(interval) %>% summarize(meanSteps=mean(steps), totalSteps=sum(steps))
print(meanPerInterval)
```

#### 3.b. Plot the Average Daily Activity Pattern
``` {r}
g = ggplot(meanPerInterval, aes(x=interval, y=meanSteps)) + geom_line(color="blue", size=1) + xlab("Interval") + ylab("Average Steps") + ggtitle("Average Steps By Interval")
print(g)
```
``` {r, echo=FALSE}
png("DailyActivityPattern.png")
dev.set(2)
dev.copy(png, "Figure/DailyActivityPattern.png")
dev.off()
```

#### 3.c. Which 5-minute interval contains the maximum steps across all days?
``` {r}
index = which(meanPerInterval$totalSteps == max(meanPerInterval$totalSteps))
meanPerInterval$interval[index]
```

# Dealing with missing values (NAs) - Computation with modified dataset
#### 4.a. Count the rows with mising values
``` {r}
print(paste("Rows with missing values", sum(is.na(activityData))))
```

#### 4.b. Create new dataset and substitue NAs with meaningful values like group mean
``` {r}
newActivityData = activityData
newActivityData = newActivityData %>% group_by(interval) %>% mutate(steps= ifelse(is.na(steps), mean(steps, na.rm=TRUE), steps))
print(head(newActivityData))
```

#### 4.c. Print the calculated data and plot the Total Steps per Day
``` {r}
newMeanPerDay = newActivityData %>% group_by(ymd(date)) %>% summarize(newMeanSteps=mean(steps), newTotalSteps = sum(steps))
print(newMeanPerDay)

g= ggplot(newMeanPerDay, aes(newTotalSteps)) + geom_histogram(fill="orange", binwidth=1000) + xlab("Steps") + ylab("Frequency") + ggtitle("Steps taken daily - NAs filled in with grouped mean")
print(g)
```
``` {r, echo=FALSE}
png("NewMeanStepsPerDay.png")
dev.set(2)
dev.copy(png, "Figure/NewMeanStepsPerDay.png")
dev.off()
```

#### 4.d. Calculate the Mean and Median of the Total steps taken
``` {r}
newMeanAndMedian = newMeanPerDay %>% na.omit() %>% summarize(NewMeanofTotalSteps = mean(newTotalSteps), NewMedianofTotalSteps = median(newTotalSteps))
print(newMeanAndMedian)
```

#### 4.e. Compare original dataset to the new dataset where the NAs are filled in with grouped median
``` {r , echo=FALSE}
      print("Original dataset")
      print(paste("Mean: ", meanAndMedian$MeanofTotalSteps, " Median: ", meanAndMedian$MedianofTotalSteps))
      print("Modified dataset - NAs filled in with group Mean")
      print(paste("Mean: ", newMeanAndMedian$NewMeanofTotalSteps, " Median: ", newMeanAndMedian$NewMedianofTotalSteps))
      print("Take away with filling in NAs:")
      print("The replacement of NAs with group mean values did not cause any major variation")
```

#### 5.a. Create new variable with two levels- "Weekday" and "Weekend" to the modified dataset
``` {r}
newActivityData$DayoftheWeek = ifelse(weekdays(as.Date(newActivityData$date)) %in% c("Monday","Tuesday","Wednesday", "Thursday", "Friday"), "Weekday", "Weekend")
print(head(newActivityData))
```

#### 5.b. Plot a time series plot of interval against weekdays/weekends
``` {r}
newMeanPerInterval = newActivityData %>% na.omit() %>% group_by(interval, DayoftheWeek) %>% summarize(newMeanSteps=mean(steps))
g = ggplot(newMeanPerInterval, aes(x=interval, y=newMeanSteps)) + geom_line(color="blue")+ facet_grid(DayoftheWeek~.) + xlab("Interval") + ylab("Average Steps") + ggtitle("Average Steps - Comparison between Weekday and Weekend")
print(g)
```
``` {r, echo=FALSE}
png("WeekdayAndWeekend.png")
dev.set(2)
dev.copy(png, "Figure/WeekdayAndWeekend.png")
dev.off()
```

      
}
=======


## Loading and preprocessing the data



## What is mean total number of steps taken per day?



## What is the average daily activity pattern?



## Imputing missing values



## Are there differences in activity patterns between weekdays and weekends?
>>>>>>> 80edf39c3bb508fee88e3394542f967dd3fd3270
